---
title: "Stock Price Prediction"
author: "Ajit Kolekar"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Get stock data for the past three years using getSymbols() fuction in quantmod package

```{r}
#Load library
library(quantmod)

# Set start date to a date three year's in the past from today
start_date <- Sys.Date() - 1095    #Three years

# Set end date to yesterday's date since the market might still be open today
end_date <- Sys.Date() - 1

# Set the symbol of the stock to analyze and predict
symbol <- "MSFT"

stock_df <- NULL

# Get stock data
stock_data <- getSymbols(Symbols = symbol, src = "yahoo", from = start_date, 
                        to = end_date, auto.assign = FALSE)
stock_data <- Cl(stock_data)

# Check the first few rows of stock data
head(stock_data)
```

Analyze the stock using technical indicators such as  Simple Moving Average, Bollinger bands, Relative Strength Index, and Moving Average Convergence Divergence 

```{r}
#stock charting
chart_Series(stock_data, col = "black")
add_SMA(n = 100, on = 1, col = "red")
add_SMA(n = 20, on = 1, col = "black")
add_RSI(n = 14, maType = "SMA")
add_BBands(n = 20, maType = "SMA", sd = 1, on = -1)
add_MACD(fast = 12, slow = 25, signal = 9, maType = "SMA", histogram = TRUE)

```

Convert the stock data into log transformation.

```{r}
### Log tranformation stock data
stock_log <- log(stock_data)
head(stock_log, n = 10)
```

Plot the log data


```{r}
plot(stock_log, main = "log stock data chart")
```

Plot auto correlation function of the log data

```{r}
acf_log <- acf(stock_log, lag.max = 500)
```

Plot partial auto correlation function of the log data

```{r}
pacf_log <- pacf(stock_log, lag.max = 500)
```

Plot the log transformed data will be differenced by 1 lag.

```{r}
### difference logged data
stock_diff <- diff(stock_log, lag = 1)

stock_diff <- na.locf(stock_diff, na.rm = TRUE,
                     fromLast = TRUE)
plot(stock_diff)
```

Split the data into training data and testing data. The last 100 days will be used as testing data. 

```{r}
### splitting into train and test data
library(caTools)
size <- length(stock_diff) - 100
train_data <- stock_diff[1:size]
```
# ARIMA Model

Build the ARIMA model

```{r}
library(forecast)
set.seed(123)
arima_model <- auto.arima(train_data, stationary = TRUE, ic = c("aicc", "aic", "bic"), 
                          trace = TRUE)

```

Display the summary details of the best ARIMA model

```{r}
summary(arima_model)
```

Diagnostic checking of the best ARIMA model

```{r}
checkresiduals(arima_model)
```

Fit the model into the training data set.

```{r}
arima <- arima(train_data, order = c(1, 0, 0))
summary(arima)

```

Forecast the data using ARIMA model

```{r}
forecast1 <- forecast(arima, h = 100)
plot(forecast1)
```

```{r}
arima <- arima(stock_log[1:size], order = c(1, 0, 0))
summary(arima)
```

Plot the forecast


```{r}
forecast_ori <- forecast(arima, h = 100)
a <- ts(stock_log)
forecast_ori %>% autoplot() + autolayer(a)
```


# GARCH Model

Model 1: Fit ARMA(0,0) GARCH(1,1) model with Student t-distribution

```{r}
library(rugarch)
garch_1 <- ugarchspec(mean.model = list(armaOrder=c(0,0)),variance.model = list(model = 'eGARCH', 
                          garchOrder = c(1, 1)),distribution = 'std')
fit_garch_1 <- ugarchfit(spec = garch_1, data= stock_diff)
fit_garch_1
```


Model 2: Fit ARMA(1,1) GARCH(1,1) model with Student t-distribution


```{r}
garch_2 <- ugarchspec(mean.model = list(armaOrder=c(1,1)),variance.model = list(model = 'eGARCH', 
                          garchOrder = c(1, 1)),distribution = 'std')

fit_garch_2 <- ugarchfit(spec = garch_2, data= stock_diff)
fit_garch_2
```

Model 3: Fit ARMA(2,2) GARCH(1,1) model with Student t-distribution


```{r}
garch_3 <- ugarchspec(mean.model = list(armaOrder=c(2,2)),variance.model = list(model = 'eGARCH', 
                          garchOrder = c(1, 1)),distribution = 'std')

fit_garch_3 <- ugarchfit(spec = garch_3, data= stock_diff)
fit_garch_3
```

Model 4: Fit ARMA(1,2) GARCH(1,1) model with Student t-distribution


```{r}
garch_4 <- ugarchspec(mean.model = list(armaOrder=c(1,2)),variance.model = list(model = 'eGARCH', 
                          garchOrder = c(1, 1)),distribution = 'std')

fit_garch_4 <- ugarchfit(spec = garch_4, data=stock_diff)
fit_garch_4
```

Model 5: Fit ARMA(2,1) GARCH(1,1) model with Student t-distribution


```{r}
garch_5 <- ugarchspec(mean.model = list(armaOrder=c(2,1)),variance.model = list(model = 'eGARCH', 
                          garchOrder = c(1, 1)),distribution = 'std')

fit_garch_5 <- ugarchfit(spec = garch_5, data=stock_diff)
fit_garch_5
```

Model 6: Fit ARMA(3,1) GARCH(1,1) model with Student t-distribution


```{r}
garch_6 <- ugarchspec(mean.model = list(armaOrder=c(3,1)),variance.model = list(model = 'eGARCH', 
                          garchOrder = c(1, 1)),distribution = 'std')

fit_garch_6 <- ugarchfit(spec = garch_6, data=stock_diff)
fit_garch_6
```

Model 7: Fit ARMA(3,2) GARCH(1,1) model with Student t-distribution


```{r}
garch_7 <- ugarchspec(mean.model = list(armaOrder=c(3,2)),variance.model = list(model = 'eGARCH', 
                          garchOrder = c(1, 1)),distribution = 'std')

fit_garch_7 <- ugarchfit(spec = garch_7, data=stock_diff)
fit_garch_7
```

Model 8: Fit ARMA(1,3) GARCH(1,1) model with Student t-distribution


```{r}
garch_8 <- ugarchspec(mean.model = list(armaOrder=c(1,3)),variance.model = list(model = 'eGARCH', 
                          garchOrder = c(1, 1)),distribution = 'std')

fit_garch_8 <- ugarchfit(spec = garch_8, data=stock_diff)
fit_garch_8
```

Find the best model 

```{r}
Model = c('fit_garch_1','fit_garch_2','fit_garch_3','fit_garch_4','fit_garch_5','fit_garch_6','fit_garch_7','fit_garch_8')
AIC = c(-5.4749, -5.4899, -5.4854, -5.4879, -5.4879, -5.4837, -5.4854, -5.4845)
(model <- data.frame(Model,AIC))

which.min(model[,'AIC'])
```
Fit the best model (model 2: ARMA(1,1) GARCH(1,1) model with Student t-distribution)

```{r}
fit_garch_2
```
Persistence of volatility

```{r}
persistence(fit_garch_2)    
```
Plot all graphs for model 2

```{r}

plot(fit_garch_2,which='all')   
```
Model convergence

```{r}
print(convergence(fit_garch_2))  
```
Forecast the data using GARCH model

```{r}
for_cast1 <-ugarchforecast(fit_garch_2,data=data,n.ahead=100)
for_cast1
```

Rolling forecast

```{r}
fit_roll <- ugarchfit(garch_5, data= stock_log[1:size], out.sample = 400)
fore_roll <- ugarchforecast(fit_roll, n.ahead=100, n.roll = 400)
fore_roll
```

Plot the rolling forecast with unconditional 1-Sigma bands

```{r}
plot(fore_roll,which=1)
```

Retrieve data and remove the reference to the stock symbol from column names and add Date as a separate column using the index. 

```{r}
start_date <- Sys.Date() - 1095    #Three years
end_date <- Sys.Date() - 1

symbol <- "MSFT"
stock_df <- NULL


getSymbols(symbol, verbose = TRUE, src = "yahoo", 
           from=start_date, to=end_date)
# stock_df = as.data.frame(get(symbol))

temp_df = as.data.frame(get(symbol))
temp_df$Date = row.names(temp_df)
row.names(temp_df) = NULL
colnames(temp_df) = c("Open", "High", "Low", "Close", 
                      "Volume", "Adjusted", "Date")
temp_df = temp_df[c("Date", "Open", "High", 
                    "Low", "Close", "Volume", "Adjusted")]
stock_df = rbind(stock_df, temp_df)

head(stock_df)
```

Save the data in CSV file for the third model LSTM to be coded using Python.

```{r}
write.csv(stock_df, 'stock_df.csv')
```
